---
title: "Week 6 Homework"
output: 
---


```{r}
#Loading needed packages
library(dplyr)
library(mice)
```



```{r}

# Loading data
cancerData <- read.csv("http://archive.ics.uci.edu/ml/machine-learning-databases/breast-cancer-wisconsin/breast-cancer-wisconsin.data", 
                       header = FALSE, 
          
                                    na.strings = "?")
#Adding column names
colnames(cancerData) <- c("SampleNo", 
                          "Thickness", 
                          "SizeUniform", 
                          "ShapeUniform", 
                          "Adhesion", 
                          "SE_CellSize", 
                          "BareNuclei", 
                          "BlandChromatin", 
                          "NormalNucleoli", 
                          "Mitoses", 
                          "Class")

```


#Question 2
1. Use the mean/mode imputation method to impute values for the missing data.
2. Use regression to impute values for the missing data.
3. Use regression with perturbation to impute values for the missing data.
4. (Optional) Compare the results and quality of classification models (e.g., SVM, KNN) build using
(1) the data sets from questions 1,2,3; (2) the data that remains after data points with missing
values are removed; and (3) the data set when a binary variable is introduced to indicate missing
values

```{r}
#Creating 3 data sets to answers questions 1-3
mmData <- cancerData
regData <- cancerData
pertData <- cancerData




```

